Consumption and resistance across states and HRRs

### 2017-02-28 states/resistance/state-report.Rmd
added parallel HRR

---
title: "Consumption and resistance across states"
bibliography: refs.bib
output:
  html_document:
    fig_height: 4
  pdf_document:
    fig_height: 3
---

@@

Here I report results of comparisons antibiotic consumption and resistance across states. States they were a convenient middle ground between census regions (in which there is certainly a consumption/resistance association, but where having only 6 data points is disappointing) and HRRs or zipcodes (in which there might be relationships, but would require throwing out a lot of consumption data because the antibiogram data is sparse).

In a separate analysis, I determined that the antibiogram data in certain locales, separating the data into those locales decreased variance more than could be expected by chance. Thus, for some bug/drug combinations, it is sensible to break the antibiogram data from census regions down into states, while for other bug/drug combinations, it does not.

@@

For each bug, consumption/resistance drug, and state, I ran a Spearman correlation test and adjusted the $p$-values using the Benjamini-Hochberg method. I also separately adjusted the $p$-values considering only the tests performed on coherent bug/drug/state combinations.

The explanatory variable in the correlation was either claims per 1,000 people per year (CPKP) or days' supply (my proxy for daily defined dose) per 1,000 people per day (DID). The dependent variable was always mean percent susceptibility of that bug to that resistance drug across antibiograms from that state.

@@

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

@@

```{r load_data}
coherence = read_tsv('../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
  rename(resistance_drug=drug) %>%
  mutate(resistance_drug=tolower(resistance_drug),
         state_coherent=level %in% c('state', 'hrr', 'zipcode')) %>%
  select(bug, resistance_drug, state_coherent)

# Antibiogram data, with individual records
abg = read_tsv('../../../../antibiogram/data/abg.tsv') %>%
  mutate(drug=tolower(drug)) %>%
  rename(resistance_drug=drug) %>%
  select(bug, resistance_drug, state, percent_susceptible)

@@

dat = bug_drug %>%
  left_join(abg, by=c('bug', 'resistance_drug')) %>%
  left_join(state_consumption, by=c('state', 'consumption_drug'))
```

@@

res = dat %>%
  group_by(bug, resistance_drug, state) %>%
  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
  left_join(state_consumption, by=c('state', 'consumption_drug')) %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(., 'mean_sus')) %>%
  ungroup %>%
  filter(!is.na(estimate)) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  mutate(record_number=1:n(),
         plot_fn=sprintf('plots/%03i.pdf', record_number),
         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
  select(bug, consumption_drug, resistance_drug, state_coherent,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig,
         plot_fn, plot_link)

coherent_res = res %>%
  filter(state_coherent) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value))
```

@@

for (i in 1:nrow(res)) {
  row = res[i,]

  plot_dat = dat %>%
    filter(bug==row$bug, consumption_drug==row$consumption_drug)

@@

```{r display_results}
res %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         state_coherent, explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

@@

```{r display_coherent_result}
coherent_res %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
     explanatory, correlation,
     adj.p.value, sig,
     plot_link) %>%
  kable
```

@@

```{r correlations_with_all_points}
dat %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(., 'percent_susceptible')) %>%
  ungroup %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  filter(!is.na(estimate), state_coherent) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  select(bug, consumption_drug, resistance_drug,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  kable
```

>>>>

---
title: "Consumption and resistance across states and HRRs"
bibliography: refs.bib
output:
  html_document:
    fig_height: 4
  pdf_document:
    fig_height: 3
---

@@

Here I report results of comparisons antibiotic consumption and resistance across states and HRRs.

In a separate analysis, I determined that the antibiogram data in certain locales, separating the data into those locales decreased variance more than could be expected by chance. Thus, for some bug/drug combinations, it is sensible to break the antibiogram data from census regions down into states, while for other bug/drug combinations, it does not. Similarly, HRRs can be coherent or not.

@@

For each bug, consumption/resistance drug, and state, I ran a Spearman correlation test and adjusted the $p$-values using the Benjamini-Hochberg method. I also separately adjusted the $p$-values considering only the tests performed on coherent bug/drug/state combinations. I performed a similar analysis using HRRs.

The explanatory variable in the correlation was either claims per 1,000 people per year (CPKP) or days' supply (my proxy for daily defined dose) per 1,000 people per day (DID). The dependent variable was always mean percent susceptibility of that bug to that resistance drug across antibiograms from that state/HRR.

@@

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, cache=TRUE,
                      echo=FALSE, warning=FALSE, message=FALSE)
```

@@

```{r load_data}
coherence = read_tsv('../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
  rename(resistance_drug=drug) %>%
  mutate(resistance_drug=tolower(resistance_drug),
         state_coherent=level %in% c('state', 'hrr', 'zipcode'),
         hrr_coherent=level %in% c('hrr', 'zipcode')) %>%
  select(bug, resistance_drug, state_coherent, hrr_coherent)

hrr = read_tsv('../../../db/hrr/hrr_2011.tsv') %>%
  select(zipcode, hrr)

# Antibiogram data, with individual records
abg = read_tsv('../../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug)) %>%
  rename(resistance_drug=drug) %>%
  left_join(hrr, by='zipcode') %>%
  select(bug, resistance_drug, state, hrr, percent_susceptible)

@@

hrr_consumption = read_tsv('../bene-consumption/hrr_usage_2011.tsv') %>%
  rename(consumption_drug=antibiotic) %>%
  select(hrr, consumption_drug, cpkp, did)

state_dat = bug_drug %>%
  left_join(abg, by=c('bug', 'resistance_drug')) %>%
  left_join(state_consumption, by=c('state', 'consumption_drug'))

hrr_dat = bug_drug %>%
  left_join(abg, by=c('bug', 'resistance_drug')) %>%
  left_join(hrr_consumption, by=c('hrr', 'consumption_drug'))
```

@@

state_res = state_dat %>%
  group_by(bug, resistance_drug, state) %>%
  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
  left_join(state_consumption, by=c('state', 'consumption_drug')) %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(., 'mean_sus')) %>%
  ungroup %>%
  filter(!is.na(estimate)) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  mutate(record_number=1:n(),
         plot_fn=sprintf('plots/%03i.pdf', record_number),
         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
  select(bug, consumption_drug, resistance_drug, state_coherent,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig,
         plot_fn, plot_link)

hrr_res = hrr_dat %>%
  group_by(bug, resistance_drug, hrr) %>%
  summarize(mean_sus=mean(percent_susceptible, na.rm=TRUE)) %>%
  left_join(bug_drug, by=c('bug', 'resistance_drug')) %>%
  left_join(hrr_consumption, by=c('hrr', 'consumption_drug')) %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(., 'mean_sus')) %>%
  ungroup %>%
  filter(!is.na(estimate)) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  mutate(record_number=1:n(),
         plot_fn=sprintf('plots/hrr-%03i.pdf', record_number),
         plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
  select(bug, consumption_drug, resistance_drug, hrr_coherent,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig,
         plot_fn, plot_link)

coherent_state_res = state_res %>%
  filter(state_coherent) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value))

coherent_hrr_res = hrr_res %>%
  filter(hrr_coherent) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value))
```

@@

for (i in 1:nrow(state_res)) {
  row = state_res[i,]

  plot_dat = state_dat %>%
    filter(bug==row$bug, consumption_drug==row$consumption_drug)

  ggsave(row$plot_fn, plot=p)
}

for (i in 1:nrow(hrr_res)) {
  row = hrr_res[i,]

  plot_dat = hrr_dat %>%
    filter(bug==row$bug, consumption_drug==row$consumption_drug)

@@

### State

```{r display_results}
state_res %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         state_coherent, explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

### HRR

```{r}
hrr_res %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
         hrr_coherent, explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

### Coherent state

@@

```{r display_coherent_result}
coherent_state_res %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
     explanatory, correlation,
     adj.p.value, sig,
     plot_link) %>%
  kable
```

### Coherent results from the HRR level

```{r}
coherent_hrr_res %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, consumption_drug,
     explanatory, correlation,
     adj.p.value, sig,
     plot_link) %>%
  kable
```

### State, all points

@@

```{r correlations_with_all_points}
state_dat %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(., 'percent_susceptible')) %>%
  ungroup %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  filter(!is.na(estimate), hrr_coherent) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  select(bug, consumption_drug, resistance_drug,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  kable
```

### HRR, all points

```{r}
hrr_dat %>%
  mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
  group_by(bug, resistance_drug, consumption_drug) %>%
  do(models(., 'percent_susceptible')) %>%
  ungroup %>%
  left_join(coherence, by=c('bug', 'resistance_drug')) %>%
  filter(!is.na(estimate), state_coherent) %>%
  mutate(adj.p.value=p.adjust(p.value, method='BH'),
         sig=pvalue_stars(adj.p.value)) %>%
  select(bug, consumption_drug, resistance_drug,
         explanatory, correlation=estimate, p.value,
         adj.p.value, sig) %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  kable
```
